---
title: "DATA"
output: html_document
date: "2025-12-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting up

Download libraries

```{r}
library("rstatix") 
library("ggplot2")
library("nlme")
library("purrr")
library("brms")
library("lme4")
library(lmerTest) 
library(multcomp)
library(dplyr)
library(tidyverse)
library(scales)
library(dplyr)
library(ggplot2)
library(cmdstanr)
library(tidyr)
library(ggeffects)

```

Download and make dataframe with all participants

```{r}
df <- read_delim("subject_000.csv",delim=";") %>%mutate(subject = "000")
df1 <- read_delim("subject_010.csv",delim=";") %>%mutate(subject = "001")
df2 <- read_delim("subject_002.csv",delim=";") %>%mutate(subject = "002")
df3 <- read_delim("subject_003.csv",delim=";") %>%mutate(subject = "003")
df4 <- read_delim("subject_004.csv",delim=";") %>%mutate(subject = "004")
df5 <- read_delim("subject_005.csv",delim=";") %>%mutate(subject = "005")
df6 <- read_delim("subject_006.csv",delim=";") %>%mutate(subject = "006")
df7 <- read_delim("subject_007.csv",delim=";") %>%mutate(subject = "007")
df8 <- read_delim("subject_008.csv",delim=";") %>%mutate(subject = "008")
df9 <- read_delim("subject_009.csv",delim=";") %>%mutate(subject = "009")
df10 <- read_delim("subject_010.csv", delim = ";") %>% mutate(subject = "010")
df11 <- read_delim("subject_011.csv", delim = ";") %>% mutate(subject = "011")
df12 <- read_delim("subject_012.csv", delim = ";") %>% mutate(subject = "012")
df13 <- read_delim("subject_013.csv", delim = ";") %>% mutate(subject = "013")
df14 <- read_delim("subject_014.csv", delim = ";") %>% mutate(subject = "014")
df15 <- read_delim("subject_011.csv", delim = ";") %>% mutate(subject = "015")

clean_subject <- function(file, subject_id) {
  read_delim(file, delim = ";") %>%
    mutate(subject = subject_id) %>%
    mutate(
      trueCat = if_else(trueCat == "narrow", 1L, 2L),
      congruentResponse = as.integer(congruentResponse)
    ) %>%
    filter(RT <= 20)
}

df  <- clean_subject("subject_000.csv", "000")
df1 <- clean_subject("subject_001.csv", "001")
df2 <- clean_subject("subject_002.csv", "002")
df3 <- clean_subject("subject_003.csv", "003")
df4 <- clean_subject("subject_004.csv", "004")
df5 <- clean_subject("subject_005.csv", "005")
df6 <- clean_subject("subject_006.csv", "006")
df7 <- clean_subject("subject_007.csv", "007")
df8 <- clean_subject("subject_008.csv", "008")
df9 <- clean_subject("subject_009.csv", "009")
df10 <- clean_subject("subject_010.csv", "010")
df11 <- clean_subject("subject_011.csv", "011")
df12 <- clean_subject("subject_012.csv", "012")
df13 <- clean_subject("subject_013.csv", "013")
df14 <- clean_subject("subject_014.csv", "014")
df15 <- clean_subject("subject_015.csv", "015")

df  <- df  %>% mutate(abs_orientation = abs(orientation))
df1 <- df1 %>% mutate(abs_orientation = abs(orientation))
df2 <- df2 %>% mutate(abs_orientation = abs(orientation))
df3 <- df3 %>% mutate(abs_orientation = abs(orientation))
df4 <- df4 %>% mutate(abs_orientation = abs(orientation))
df5 <- df5 %>% mutate(abs_orientation = abs(orientation))
df6 <- df6 %>% mutate(abs_orientation = abs(orientation))
df7 <- df7 %>% mutate(abs_orientation = abs(orientation))
df8 <- df8 %>% mutate(abs_orientation = abs(orientation))
df9 <- df9 %>% mutate(abs_orientation = abs(orientation))
df10 <- df10 %>% mutate(abs_orientation = abs(orientation))
df11 <- df11 %>% mutate(abs_orientation = abs(orientation))
df12 <- df12 %>% mutate(abs_orientation = abs(orientation))
df13 <- df13 %>% mutate(abs_orientation = abs(orientation))
df14 <- df14 %>% mutate(abs_orientation = abs(orientation))
df15 <- df15 %>% mutate(abs_orientation = abs(orientation))


data <- dplyr::bind_rows(
  df, df1, df2, df3, df4, df5, df6, df7,
  df8, df9, df10, df11, df12, df13, df14, df15
)
```

# Behavioral analyses

## Contrast levels

```{r}
#Accuracy x contrast
acc_contrast <- data %>%
  group_by(contrast) %>%
  summarise(
    n = n(),
    p_corr = mean(congruentResponse),
    se = sd(congruentResponse) / sqrt(n)
  )


ggplot(acc_contrast,
                   aes(x = contrast, y = p_corr)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = p_corr - se, ymax = p_corr + se),
                width = 0.005) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Contrast level (%)",
       y = "Proportion correct (%)")+
    theme_bw(base_size = 14)



#confidence x contrast
conf_contrast <- data %>%
  group_by(contrast) %>%              
  summarise(
    n       = n(),
    m_conf  = mean(confRating, na.rm = TRUE),             
    se_conf = sd(confRating, na.rm = TRUE) / sqrt(n)       
  )

ggplot(conf_contrast,
                   aes(x = contrast, y = m_conf)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = m_conf - se_conf, ymax = m_conf + se_conf),
                width = 0.005) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Contrast level (%)",
       y = "Mean Confidence")+
    theme_bw(base_size = 14)

#Rt x contrast
rt_contrast <- data %>%
  group_by(contrast) %>%  
  summarise(
    n       = n(),
    mean_rt = mean(RT, na.rm = TRUE),
    se_rt   = sd(RT, na.rm = TRUE) / sqrt(n)
  )

ggplot(rt_contrast,
                aes(x = contrast, y = mean_rt)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_rt - se_rt, ymax = mean_rt + se_rt),width = 0.005) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Contrast level (%)",
       y = "Reaction time (s)") +
  theme_bw(base_size = 14)

```

## Orientation

```{r}
data <- data %>%
  mutate(
    orientation_bin = cut(
      abs_orientation,
      breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40),
      include.lowest = TRUE,
      right = FALSE,
      labels = c(
        "0–5", "5–10", "10–15", "15–20",
        "20–25", "25–30", "30–35", "35–40"
      )
    )
  )

# Accuracy x bin
acc_bin <- data %>%
  group_by(orientation_bin) %>%
  summarise(
    n      = n(),
    p_corr = mean(congruentResponse),
    se     = sd(congruentResponse) / sqrt(n),
    .groups = "drop"
  ) %>%
  filter(!is.na(orientation_bin))     # fjern NA-bin

ggplot(acc_bin, aes(x = orientation_bin, y = p_corr, group = 1)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = p_corr - se,
                    ymax = p_corr + se),
                width = 0.1) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  labs(
    x = "Orientation bin (°)",
    y = "Porportion correct (%)"
  ) +
  theme_bw(base_size = 14)

# Mean confidence x orientation
conf_bin <- data %>%
  group_by(orientation_bin) %>%
  summarise(
    n        = n(),
    mean_conf = mean(confRating, na.rm = TRUE),
    se_conf   = sd(confRating, na.rm = TRUE) / sqrt(n)
  )%>%
  filter(!is.na(orientation_bin))     # fjern NA-bin

ggplot(conf_bin, aes(x = orientation_bin, y = mean_conf, group = 1)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_conf - se_conf,
                    ymax = mean_conf + se_conf),
                width = 0.1) +
  labs(
    x = "Orientation bin (°)",
    y = "Mean confidence"
  ) +
  theme_bw(base_size = 14)

# RT x orientation 
rt_bin <- data %>%
  group_by(orientation_bin) %>%
  summarise(
    n       = n(),
    mean_rt = mean(RT, na.rm = TRUE),
    se_rt   = sd(RT, na.rm = TRUE) / sqrt(n)
  )%>%
  filter(!is.na(orientation_bin)) 
  
ggplot(rt_bin, aes(x = orientation_bin, y = mean_rt, group = 1)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_rt - se_rt,
                    ymax = mean_rt + se_rt),
                width = 0.1) +
  labs(
    x = "Orientation bin (°)",
    y = "Reaction time (s)"
  ) +
  theme_bw(base_size = 14)

```

# Modelling

Make stan lists for all participants

```{r}
stan_data <- list(
  N = nrow(df),
  s = df$abs_orientation,
  c = df$contrast,
  response = as.integer(df$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data1 <- list(
  N = nrow(df1),
  s = df1$abs_orientation,
  c = df1$contrast,
  response = as.integer(df1$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data2 <- list(
  N = nrow(df2),
  s = df2$abs_orientation,
  c = df2$contrast,
  response = as.integer(df2$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data3 <- list(
  N = nrow(df3),
  s = df3$abs_orientation,
  c = df3$contrast,
  response = as.integer(df3$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data4 <- list(
  N = nrow(df4),
  s = df4$abs_orientation,
  c = df4$contrast,
  response = as.integer(df4$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data5 <- list(
  N = nrow(df5),
  s = df5$abs_orientation,
  c = df5$contrast,
  response = as.integer(df5$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data6 <- list(
  N = nrow(df6),
  s = df6$abs_orientation,
  c = df6$contrast,
  response = as.integer(df6$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data7 <- list(
  N = nrow(df7),
  s = df7$abs_orientation,
  c = df7$contrast,
  response = as.integer(df7$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data8 <- list(
  N = nrow(df8),
  s = df8$abs_orientation,
  c = df8$contrast,
  response = as.integer(df8$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data9 <- list(
  N = nrow(df9),
  s = df9$abs_orientation,
  c = df9$contrast,
  response = as.integer(df9$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data10 <- list(
  N = nrow(df10),
  s = df10$abs_orientation,
  c = df10$contrast,
  response = as.integer(df10$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data11 <- list(
  N = nrow(df11),
  s = df11$abs_orientation,
  c = df11$contrast,
  response = as.integer(df11$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data12 <- list(
  N = nrow(df12),
  s = df12$abs_orientation,
  c = df12$contrast,
  response = as.integer(df12$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data13 <- list(
  N = nrow(df13),
  s = df13$abs_orientation,
  c = df13$contrast,
  response = as.integer(df13$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data14 <- list(
  N = nrow(df14),
  s = df14$abs_orientation,
  c = df14$contrast,
  response = as.integer(df14$responseCat),
  sigma1 = 3,
  sigma2 = 12
)

stan_data15 <- list(
  N = nrow(df15),
  s = df15$abs_orientation,
  c = df15$contrast,
  response = as.integer(df15$responseCat),
  sigma1 = 3,
  sigma2 = 12
)




```

Collect lists into one big list

```{r}
stan_list <- list(
  "000" = stan_data,
  "001" = stan_data1,
  "002" = stan_data2,
  "003" = stan_data3,
  "004" = stan_data4,
  "005" = stan_data5,
  "006" = stan_data6,
  "007" = stan_data7,
  "008" = stan_data8,
  "009" = stan_data9,
  "010" = stan_data10,
  "011" = stan_data11,
  "012" = stan_data12,
  "013" = stan_data13,
  "014" = stan_data14,
  "015" = stan_data15
)


```

Run MLE

```{r}
model <- cmdstan_model("mle2.stan")

mle_params <- imap_dfr(stan_list, function(d, sid) {
  fit <- model$optimize(data = d, seed = 1)
  est <- fit$mle()

  tibble(
    subject = sid,
    N = d$N,
    alpha  = est["alpha"],
    beta   = est["beta"],
    gamma  = est["gamma"],
    lambda = est["lambda"],
  )
})

mle_params
```

```{r}
seeds <- c(123, 456, 789, 2024)

fits <- lapply(seeds, function(s) {
  cmdstanr::cmdstan_model("mle2.stan")$optimize(
    data = stan_data,
    seed = s
  )
})
```

Plot K(σ)

```{r}
# Defining functions
sigma_from_contrast <- function(c, alpha, beta, gamma) {
  sqrt((alpha * c)^(-beta) + gamma)
}

k_from_sigma <- function(sigma, sigma1 = 3, sigma2 = 12) {
  k1 <- 0.5 * log((sigma^2 + sigma2^2) / (sigma^2 + sigma1^2)) 
  k2 <- (sigma2^2 - sigma1^2) / (2 * (sigma^2 + sigma1^2) * (sigma^2 + sigma2^2))
  sqrt(k1 / k2)
}

# contrast levels
contrast_levels <- sort(unique(data$contrast))


# k as a function pr subject x contrast
k_by_subject_contrast <- expand_grid(
  subject = unique(mle_params$subject),
  contrast = sort(unique(data$contrast))
) %>%
  left_join(mle_params, by = "subject") %>%
  mutate(
    sigma = sigma_from_contrast(contrast, alpha, beta, gamma),
    k = k_from_sigma(sigma, sigma1 = 3, sigma2 = 12)
  )

# errorbars 
k_summary <- k_by_subject_contrast %>%
  group_by(contrast) %>%
  summarise(
    mean_k = mean(k, na.rm = TRUE),
    sd_k   = sd(k, na.rm = TRUE),
    n      = sum(!is.na(k)),
    se_k   = sd_k / sqrt(n),
    .groups = "drop"
  )



```

Plot

```{r}
# plot
ggplot(k_summary, aes(x = factor(contrast), y = mean_k)) +
  geom_line(size=3)+
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = mean_k - se_k, ymax = mean_k + se_k),
    width = 0.1
  ) +
  geom_line(aes(group = 3)) +
  scale_x_discrete(
    labels = function(x) paste0(round(as.numeric(x) * 100, 1), "%")
  ) +
  labs(
    x = "Contrast",
    y = "K(σ)",
    title = "K(σ) as a function of contrast (mean ± SEM)"
  ) +
  theme_bw()

```

```{r}
df_plot <- k_by_subject_contrast %>%
  mutate(
    subject  = factor(subject),
    contrast = factor(contrast),   # treat contrast as discrete
    k        = as.numeric(k)
  )

ggplot(df_plot, aes(x = factor(contrast), y = k, group = subject, color = subject)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +
  geom_point(size = 2, alpha = 0.8) +
  labs(
    x = "Contrast",
    y = "K(σ)",
    color = "Participant",
    title = "K(σ) pr. participant as a function of contrast"
  ) +
   scale_x_discrete(
    labels = function(x) paste0(round(as.numeric(x) * 100, 1), "%")
  )+
  theme_bw() +
  theme(
    legend.position = "right",
  )


```

# Confidence analyses

Does decision boundary systemtically change with contrast

```         
```

When contrast is higher –\> decision boundary is significantly lower

Does distance from K(sigma) say somehing about

```{r}
#Calculate Distances from s_i to K(sigma_i) and add to 

data_with_k <- data %>%
  left_join(
    k_by_subject_contrast %>% select(subject, contrast, k, sigma),
    by = c("subject", "contrast")
  )
data_with_k <- data_with_k %>%
  mutate(
    D = abs(abs(orientation) - k)
  )

data_with_k <- data_with_k %>%
  mutate(D_star = k - abs(orientation))


data_with_k$congruentResponse <- factor(
  data_with_k$congruentResponse,
  levels = c(0, 1),
  labels = c("Incongruent", "Congruent")
)

data_with_k <- data_with_k %>%
  group_by(contrast) %>%
  mutate(
    D_within = D - mean(D)
  ) %>%
  ungroup()


distance_lm <- lmer(
  confRating ~ D_within + congruentResponse + (1 | subject),
  data = data_with_k
)

summary(distance_lm)

pD <- ggpredict(
  distance_lm,
  terms = "D_within [all]"
)

ggplot(pD, aes(x = x, y = predicted)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.25) +
  labs(
    x = "Distance from decision boundary K(σ) (D)",
    y = "Predicted confidence",
    title = "Effect of distance from decision boundary on confidence"
  ) +
  theme_bw(base_size = 14)

pCR <- ggpredict(
  distance_lm,
  terms = "congruentResponse"
)

ggplot(pCR, aes(x = x, y = predicted)) +
  geom_point(size = 4) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.15
  ) +
  labs(
    x = "Response congruency",
    y = "Predicted confidence",
    title = "Effect of response congruency on confidence"
  ) +
  theme_bw(base_size = 14)
```

```{r}


# Vælg antal bins (12–20 er ofte fint)
n_bins <- 30

max_abs <- max(abs(data_with_k$D_within), na.rm = TRUE)
breaks_sym <- seq(-max_abs, max_abs, length.out = n_bins + 1)

df_binned_signed <- data_with_k %>%
  mutate(
    D_bin = cut(D_within, breaks = breaks_sym, include.lowest = TRUE)
  ) %>%
  group_by(D_bin) %>%
  summarise(
    D_mid = mean(D_within, na.rm = TRUE),
    mean_conf = mean(confRating, na.rm = TRUE),
    se_conf = sd(confRating, na.rm = TRUE) / sqrt(sum(!is.na(confRating))),
    n = sum(!is.na(confRating)),
    .groups = "drop"
  )

ggplot(df_binned_signed, aes(x = D_mid, y = mean_conf)) +
  geom_hline(yintercept = mean(data_with_k$confRating, na.rm = TRUE),
             linetype = "dashed", linewidth = 0.7, alpha = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.7, alpha = 0.6) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_conf - se_conf, ymax = mean_conf + se_conf),
                width = 0) +
  labs(
    x = expression(paste("Within-contrast centered distance (", hat(delta)["within"], ")")),
    y = "Mean confidence (± SE)",
    title = "Confidence as a function of within-contrast distance"
  ) +
  theme_bw(base_size = 14)

```
